---
title: "Roulette Problem"
author: 'Jiayu Shi'
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    code_folding: hide
    toc: yes
    number_sections: yes
    toc_depth: 3
    toc_float: yes
---

```{r global options, include = FALSE}
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE, cache=TRUE)
```


# Introduction 

In this blog, I will focus on one strategy used in Roulette. A roulette table composed of 38 evenly sized pockets on a wheel. The pockets are colored red, black, or green. Roulette is a game of chance in which a pocket is randomly selected. Gamblers may place a wager that the randomly selected pocket will be red or other colors. The payout for a bet on black (or red) is \$1 for each \$1 wagered. This means that if a gambler bets \$1 on black and the randomly selected pocket is black, then the gambler will get the original \$1 wager and an additional \$1 as winnings.


# Methods

Below this the code I use to simulate the Roulette.

The first part of the code is to define the functions I will use later.
```{r}
single_spin <- function(){
  possible_outcomes <- c(rep("red",18), rep("black",18), rep("green",2))
  sample(possible_outcomes, 1)
}

martingale_wager <- function(
  previous_wager
  , previous_outcome
  , max_wager
  , current_budget
){
  if(previous_outcome == "red") return(1)
  min(2*previous_wager, max_wager, current_budget) 
}

one_play <- function(previous_ledger_entry, max_wager){
  # Create a copy of the input object that will become the output object
  out <- previous_ledger_entry
  out[1, "game_index"] <- previous_ledger_entry[1, "game_index"] + 1
  out[1, "starting_budget"] <- previous_ledger_entry[1, "ending_budget"]
  out[1, "wager"] <- martingale_wager(
    previous_wager = previous_ledger_entry[1, "wager"]
    , previous_outcome = previous_ledger_entry[1, "outcome"]
    , max_wager = max_wager
    , current_budget = out[1, "starting_budget"]
  )
  out[1, "outcome"] <- single_spin()
  out[1, "ending_budget"] <- out[1, "starting_budget"] + 
    ifelse(out[1, "outcome"] == "red", +1, -1)*out[1, "wager"]
  return(out)
}

one_series <- function(
  max_games, starting_budget, winning_threshold, max_wager
){
  # Initialize ledger
  ledger <- data.frame(
      game_index = 0:max_games
    , starting_budget = NA_integer_
    , wager = NA_integer_
    , outcome = NA_character_
    , ending_budget = NA_integer_
  )
  ledger[1, "wager"] <- 1
  ledger[1, "outcome"] <- "red"
  ledger[1, "ending_budget"] <- starting_budget
  for(i in 2:nrow(ledger)){
    #browser()
    ledger[i,] <- one_play(ledger[i-1,], max_wager)
    if(stopping_rule(ledger[i,], winning_threshold)) break
  }
  # Return non-empty portion of ledger
  ledger[2:i, ]
}

stopping_rule <- function(
  ledger_entry
  , winning_threshold
){
  ending_budget <- ledger_entry[1, "ending_budget"]
  if(ending_budget <= 0) return(TRUE)
  if(ending_budget >= winning_threshold) return(TRUE)
  FALSE
}

profit <- function(ledger){
  n <- nrow(ledger)
  profit <- ledger[n, "ending_budget"] - ledger[1, "starting_budget"]
  return(profit)
}

```

The second part is aimed to generate 2 examples of different gamblers.
The first gamblers loses the game, while the second one win the game.
```{r}
#require(magrittr)

#svg(filename = "loser.svg", width=16, height =9)
par(cex.axis=2, cex.lab = 2, mar = c(8,8,2,2), bg = rgb(222, 235, 247, max = 255))
set.seed(1)
ledger <- one_series(200,200,300,500)
plot(ledger[,c(1,5)], type = "l", lwd = 5, xlab = "Game Index", ylab = "Budget")
#dev.off()

#svg(filename = "winner.svg", width=16, height =9)
par(cex.axis=2, cex.lab = 2, mar = c(8,8,2,2), bg = rgb(222, 235, 247, max = 255))
set.seed(2)
l2 <- one_series(200,200,300,500)
plot(l2[,c(1,5)], type = "l", lwd = 5, xlab = "Game Index", ylab = "Budget")
#dev.off()
```

# Results
In this part, I will try to find the average profit of the Roulette.
Below is my code to calculate the average.
```{r}
#initialParameter = c(200,200,300,500)
set.seed(1)
profitlist = replicate(100,profit(one_series(200,200,300,500)))
plot(profitlist)
(averageProfit = mean(profitlist))

```
And I find the average of 100 rounds of Roulette is -21.29.
# Conclusions
Because the average of 100 rounds of Roulette is -21.29, the martingale strategy has been proven false. The reason why this strategy does not work is that there are limits which constraint gamblers to continue infinitely, such as the limited starting budget, the max wager.


